An automatic approach to find all possible calendar effects (periodic patterns) for an univariate time series

### 2021-02-21 paper/main.Rmd
Added lm and glm results

## Definition of the distance measure

@@

## Properties of the distance measure

@@

The distribution of wpd is different for different levels of facets and x-axis levels. This is because the statistic maximum which is used to define $wpd$ is affected by the number of categories. The measure would have higher values if $A$ or $B$ has higher levels. However, we would ideally want a higher value of the measure only if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories $J$ or $K$ is high. Therefore, in order to compare $wpd$ across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure. Henceforth we call the measure already discussed as $wpd_{raw}$ and the normalized measure as $wpd_{norm}$. The measure $wpd_{norm}$ could potentially lead to comparison of the measure across different panels and also identifying only the interesting panels from a data set. We discuss two approaches for normalization, both of which are based on the simulation results conducted in the earlier section.

@@

The permutation approach ensures that the distribution of the normalized distance measure has the same mean and standard deviation across all combinations of $nx_i$ and $nfacet_j$. The normalized distances through permutation is computed as follows: $x^{perm}_{k} =  (x_{k} - mean_l(x_{k,l}))/sd_l(x_{k, l})$,  $x^{perm}_{k}$ is the value of the $wpd_{perm$ for the $k^{th}$ panel. Standardizing the variable $wpd_{perm}$ in this approach leads to $location = 0$ and $scale = 1$ for this variable.

@@

@@

```{r linear-model}
G21 <- read_rds(here("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_Gamma21.rds"))

@@

@@

The tabulated values and graphical representations of the simulation results are provided in Appendix. The learning from the simulations are as follows:
The values of the measure $wpd_{raw}$ is least for $D_{null}$, followed by 
$D_{var_f}$, $D_{var_x}$) and $D_{var_{all}}$). This is a desirable result since the measure $wpd_{raw}$ was designed such that this relationship holds.
Furthermore, the distribution of the measure $wpd_{raw}$ changes for different facet and x categories. The location of the distribution shifts to the right and it also becomes more skewed for more higher facet and x-axis categories. Now this is not desirable, as it would mean that we can't compare the $wpd_{raw}$ values across different panels. The distribution of $wpd_{norm}$ looks more similar with at least the mean and standard of the distributions being uniform across panels. This means,  $wpd_{norm}$ could be used to measure differences in distribution across panels. Also, note that since the data is processed using normal-quantile-transform, this measure is independent of the intial distribution of the underlying data and hence is also comparable across different data sets. This is valid for the case when sample size $ntimes$ for each combination of categories is at least 30 and  $nperm$ used for computing $wpd_{norm}$ is at least 100. More detailed results about the properties of $wpd_{raw}$ and $wpd_{norm}$ could be found in Appendix.

>>>>

## Definition

@@

## Properties

@@

The distribution of wpd is different for different levels of facets and x-axis levels. This is because the statistic maximum which is used to define $wpd$ is affected by the number of categories. The measure would have higher values if $A$ or $B$ has higher levels. However, we would ideally want a higher value of the measure only if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories $J$ or $K$ is high. Therefore, in order to compare $wpd$ across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure. Henceforth we call the measure already discussed as $wpd_{raw}$ and the normalized measure as $wpd_{norm}$. The measure $wpd_{norm}$ could potentially lead to comparison of the measure across different panels and also identifying only the interesting panels from a data set. We discuss two approaches for normalization, both of which are based on the simulation results.

@@

The permutation approach ensures that the distribution of the normalized distance measure has the same mean and standard deviation across all combinations of $nx_i$ and $nfacet_j$. The normalized distances through permutation is computed as follows: $x^{perm}_{k} =  (x_{k} - mean_l(x_{k,l}))/sd_l(x_{k, l})$,  $x^{perm}_{k}$ is the value of the $wpd_{perm}$ for the $k^{th}$ panel. Standardizing the variable $wpd_{perm}$ in this approach leads to $location = 0$ and $scale = 1$ for this variable.

@@

_Linear model_

@@

```{r linear-model,message=FALSE}
G21 <- read_rds(here("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_Gamma21.rds"))

@@

_Generalised linear model_

In the linear model approach, $wpd_{raw} \in R$ was assumed, whereas, $wpd_{raw}$, Jensen-Shannon Distance (JSD) lies between 0 and 1. Furthermore, JSD follows a Chi-square distribution, which is a special case of Gamma distribution and hence belongs to exponential family of distributions. Therefore, we can fit a generalized linear model instead of a linear model to allow for the response variable to follow a Gamma distribution. The inverse link is used when we know that the mean response is bounded, which is applicable in our case since $0 \leq wpd_{raw}\leq 1$.

We fit a Gamma generalized linear model with the inverse link which is of the form:  
$$y = a+b*log(x) + e$$, where $y = median(wpd_{raw})$, $x = nx*nfacet$.
Let $E(y) = \mu$ and $a + b*log(x) = g(\mu)$ where $g$ is the link function. Then $g(\mu)=  1/mu$ and $\hat \mu  = 1/(\hat a + \hat b log(x))$. The residuals from this model $(y-\hat y) = (y-1/(\hat a + \hat b log(x)))$ would be expected to have no dependency on $x$. Thus,  $wpd_{glm}$ is chosen as the residuals from this model and is defined as:
$wpd_{glm} = wpd_{raw} - 1/(\hat a + \hat b*log(nx*nfacet))$.

<!-- Please note: -->

<!-- - subtracting the intercept brings the location of the transformed variable to 0 for either case -->
<!-- - heterogeneity is higher for this case after normalization as compared to the earlier one. -->

```{r wpd-glm-horizontal, echo = FALSE, message=FALSE, fig.caption = "$wpd_{glm}$ is plotted against log(nx*nfacet) and this transformation leads to the median of $wpd_{glm}$ to have no relationship with log(nx*nfacet)"}

G21 <- read_rds(here("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_Gamma21.rds"))

glm_fit <- glm(actual ~ log(`nx * nfacet`),
              family = Gamma(link = "inverse"),
              data = G21_median)

intercept <- glm_fit$coefficients[1]
slope <- glm_fit$coefficients[2]

G21_sd  = G21 %>% 
    mutate(wpd_glm =  (value - (1/(intercept + slope*log(nx*nfacet)
    )
    )
    ))

scale_fac <- 1/G21_sd$wpd_glm %>% sd()

# checking the fit of the residuals from glm fit

# fitted_glm <- fitted(glm_fit, type = "response")
# residuals <- residuals.glm(glm_fit, type = "response")
# hist(residuals)

#h = augment(glm_fit)
# ggplot(h) +
#   geom_histogram(aes(x = .resid))

# residual <- G21_median$actual  - fitted_glm
# 
# hist(residual)

G21 %>% 
  ggplot(aes(x=log(nx*nfacet),
             y = (value - (1/(intercept + slope*log(nx*nfacet)
                    )
                    )
             )
  )
  ) +
  geom_point() + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + 
  ylab("wpd_glm = wpd_raw - 1/(a  + b*log(nx*nfacet))")
```

```{r wpd-glm-dist, fig.cap = " The distribution of $wpd_{glm}$ is plotted. The distributions are more similar across higher nx and nfacet and dissimilar for fewer nc and nfacets."}

G21_glm <- G21 %>% 
  mutate(wpd_glm =  (value - (1/(intercept + slope*log(nx*nfacet)
                    )
                    )
             ),
         wpd_glm_scaled = ((wpd_glm*320)))

G21_glm$wpd_glm_scaled %>% sd()

G21_glm %>% 
  ggplot() +
  geom_density(aes(x = wpd_glm), 
                   fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") +
  theme(legend.position = "bottom") 
```

@@

The tabulated values and graphical representations of the simulation results are provided in Appendix. The learning from the simulations are as follows:
The values of the measure $wpd_{raw}$ is least for $D_{null}$, followed by 
$D_{var_f}$, $D_{var_x}$) and $D_{var_{all}}$). This is a desirable result since the measure $wpd_{raw}$ was designed such that this relationship holds.
Furthermore, the distribution of the measure $wpd_{raw}$ changes for different facet and x categories. The location of the distribution shifts to the right and it also becomes more skewed for more higher facet and x-axis categories. Now this is not desirable, as it would mean that we can't compare the $wpd_{raw}$ values across different panels. The distribution of $wpd_{norm}$ looks more similar with at least the mean and standard of the distributions being uniform across panels. This means,  $wpd_{norm}$ could be used to measure differences in distribution across panels. Also, note that since the data is processed using normal-quantile-transform, this measure is independent of the initial distribution of the underlying data and hence is also comparable across different data sets. This is valid for the case when sample size $ntimes$ for each combination of categories is at least 30 and  $nperm$ used for computing $wpd_{norm}$ is at least 100. More detailed results about the properties of $wpd_{raw}$ and $wpd_{norm}$ could be found in Appendix.
